\texttt{davos}: a Python package ``smuggler'' for constructing lightweight reproducible notebooks

### 2022-11-03 paper/main.tex
wording tweaks to section 3

After importing \texttt{davos} (line 1), the user first smuggles two utilities for checking and reading local files in the code below. The \texttt{smuggle} statement in line 3 loads the \texttt{is\_file()} function from the Python standard library's \texttt{os.path} module. Standard library modules are included with all Python distributions, so this line is functionally equivalent to an \texttt{import} statement and no onion comment is needed. Line 4 loads the \texttt{joblib} library~\cite{Varo10}, installing it first, if necessary. Since \texttt{joblib}'s I/O interface has historically remained stable and backwards-compatible across releases, requiring that users have a particular exact version installed would likely be unnecessarily restrictive. However, there's also no guarantee that a \textit{future} release won't introduce some breaking change, so to help ensure the analysis notebook continues to run properly in the future, the onion comment in line 4 limits allowable versions to those already released when the code was written.
%After importing \texttt{davos} (line 1), the user first smuggles two utilities that will be used to help find and load local files in the code below: the \texttt{is\_file} function from the \texttt{os.path} module, and the \texttt{joblib} library~\cite{Varo10} (lines 3--4). The first \texttt{smuggle} statement needs no onion comment, as the \texttt{os.path} module is part of Python's standard library and is therefore guaranteed to be available. The second ensures \texttt{joblib} is installed before loading it Enforcing an exact version of \texttt{joblib} is likely unnecessary, since the package's I/O interface has historically remained generally stable across versions. However, there's no

Line 6 then uses the \texttt{davos.config} object to enable \texttt{davos}'s \texttt{auto\_rerun} option before smuggling the next two packages, \texttt{NumPy}~\cite{HarrEtal20} and \texttt{pandas}. Because these packages rely heavily on custom C data types, loading them may require restarting the notebook kernel if different versions were already imported during the same interpreter session (see Section~\ref{subsec:config}). This is unlikely to happen with \texttt{NumPy}, as it is loaded explicitly only once (by line 7) and is not required by any packages loaded before it (and therefore not implicitly imported when loading them). However, it's possible a future user could edit this notebook to add additional code before line 7, run the notebook's cells out of order, or even configure their IPython kernel to execute custom code on startup (e.g., as Google Colab notebooks do), any of which could potentially import a different, existing version of \texttt{NumPy} before \texttt{davos} smuggles the version specified in the onion comment. Enabling the \texttt{auto\_rerun} option before smuggling \texttt{NumPy} would simplify running the notebook in these scenarios, but the main benefit it affords comes when smuggling the \texttt{pandas} library.

The dataset that the user wants to analyze was provided as a \texttt{pandas.Panel} object, serialized with Python's built-in \texttt{pickle} module. The \texttt{pickle} protocol is a popular choice for data persistence in Python, allowing users to save, share, and load arbitrary objects, though with an important caveat: A ``pickled'' object's class definition must be importable from the same module as when it was saved in order to successfully ``unpickle'' (i.e., load and restore) it later. The \texttt{Panel} class was removed from \texttt{pandas} in July, 2019 (v0.25.0), meaning that the user's dataset can be loaded only if their notebook uses a \texttt{pandas} version released prior to that change. However, updates to the \texttt{pandas} library since then have brought substantial improvements that the user would like to take advantage of in their analysis code, including faster performance, more memory-efficient data types, new functions and methods, compatibility with newer versions of Python and popular plotting libraries, and better display formats (including in Jupyter notebooks).
%Recent updates to the \texttt{pandas} library have brought substantial improvements that the user would like to take advantage of in their analysis code, including faster performance, more memory-efficient data types, new functions and methods, compatibility with newer versions of Python and popular plotting libraries, and better display formats (including in Jupyter notebooks). However, the \texttt{Panel} class was removed from \texttt{pandas} in July, 2019 (v0.25.0), meaning that the dataset they are working with can only be loaded by a version released prior to that change.
\texttt{davos} makes it possible to do both within a single notebook, and with the \texttt{auto\_rerun} option, this process can be fully automated. The first time the notebook is run, lines 10--12 will install and load an older \texttt{pandas} version that defines the \texttt{Panel} class, read the dataset from its serialized format, convert it from a \texttt{Panel} to a multi-index \texttt{DataFrame}, and write its contents to a CSV file. The \texttt{smuggle} statement on line 14 then installs a newer \texttt{pandas} version (v1.3.5) for use in the analyses themselves. Because an earlier version of \texttt{pandas} was previously loaded (by line 10), and because the package's C extension modules were updated between the old and new versions, switching to the new version will require restarting the notebook kernel. With the \texttt{auto\_rerun} option enabled, \texttt{davos} will handle this automatically, and once the kernel is restarted, will re-run the notebook up through the current cell (as well as re-queue any other cells that were queued before restarting). This second time the notebook is run (as well as any future times),

>>>>

After importing \texttt{davos} (line 1), the user first smuggles two utilities for interacting with local files in the code below. The \texttt{smuggle} statement in line 3 loads the \texttt{is\_file()} function from the Python standard library's \texttt{os.path} module. Standard library modules are included with all Python distributions, so this line is functionally equivalent to an \texttt{import} statement and no onion comment is needed. Line 4 loads the \texttt{joblib} library~\cite{Varo10}, installing it first, if necessary. Since \texttt{joblib}'s I/O interface has historically remained stable and backwards-compatible across releases, requiring that users have a particular exact version installed would likely be unnecessarily restrictive. However, there's also no guarantee that a \textit{future} release won't introduce some breaking change, so to help ensure the analysis notebook continues to run properly in the future, the onion comment in line 4 limits allowable versions to those already released when the code was written.
%After importing \texttt{davos} (line 1), the user first smuggles two utilities that will be used to help find and load local files in the code below: the \texttt{is\_file} function from the \texttt{os.path} module, and the \texttt{joblib} library~\cite{Varo10} (lines 3--4). The first \texttt{smuggle} statement needs no onion comment, as the \texttt{os.path} module is part of Python's standard library and is therefore guaranteed to be available. The second ensures \texttt{joblib} is installed before loading it Enforcing an exact version of \texttt{joblib} is likely unnecessary, since the package's I/O interface has historically remained generally stable across versions. However, there's no

Line 6 then uses the \texttt{davos.config} object to enable \texttt{davos}'s \texttt{auto\_rerun} option before smuggling the next two packages, \texttt{NumPy}~\cite{HarrEtal20} and \texttt{pandas}. Because these packages rely heavily on custom C data types, loading them may require restarting the notebook kernel if different versions were already imported during the same interpreter session (see Section~\ref{subsec:config}). This is unlikely to happen with \texttt{NumPy}, as it is loaded explicitly only once (by line 7) and is not required by any packages loaded before it (and therefore not implicitly imported while loading them). However, it's possible a future user could edit this notebook to add additional code before line 7, run the notebook's cells out of order, or even configure their IPython kernel to execute custom code on startup (e.g., as Google Colab notebooks do), any of which could potentially import a different, existing version of \texttt{NumPy} before \texttt{davos} smuggles the version specified in the onion comment. Enabling the \texttt{auto\_rerun} option before smuggling \texttt{NumPy} helps to simplify running the notebook in these scenarios, but the main benefit it affords comes when smuggling the \texttt{pandas} library.

The dataset that the user wants to analyze was provided as a \texttt{pandas.Panel} object, serialized with Python's built-in \texttt{pickle} module. The \texttt{pickle} protocol is a popular choice for data persistence in Python, allowing users to save, share, and load arbitrary objects, though with an important caveat: A ``pickled'' object's class definition must be importable from the same module as when it was saved in order to successfully ``unpickle'' (i.e., load and restore) it later. The \texttt{Panel} class was removed from \texttt{pandas} in July, 2019 (v0.25.0), meaning that the user's dataset can be loaded only if their notebook uses a \texttt{pandas} version released prior to that change. However, updates to the \texttt{pandas} library since then have brought substantial improvements that the user would like to take advantage of in their analysis code, including faster performance, more memory-efficient data types, new functions and methods, compatibility with newer versions of Python and popular plotting libraries, and better display formats (including in Jupyter notebooks).
%Recent updates to the \texttt{pandas} library have brought substantial improvements that the user would like to take advantage of in their analysis code, including faster performance, more memory-efficient data types, new functions and methods, compatibility with newer versions of Python and popular plotting libraries, and better display formats (including in Jupyter notebooks). However, the \texttt{Panel} class was removed from \texttt{pandas} in July, 2019 (v0.25.0), meaning that the dataset they are working with can only be loaded by a version released prior to that change.
\texttt{davos} makes it possible to do both within a single notebook, and with the \texttt{auto\_rerun} option, this process can be fully automated. The first time the notebook is run, lines 10--12 will install and load an older \texttt{pandas} version that defines the \texttt{Panel} class, use it to read the dataset from its serialized file format, convert the dataset from a \texttt{Panel} to a multi-index \texttt{DataFrame}, and write its contents to a CSV file. The \texttt{smuggle} statement on line 14 then installs a newer version of \texttt{pandas} (v1.3.5) for use in the analyses themselves. Because an older \texttt{pandas} version was previously loaded (by line 10), and because the package's C extension modules were modified between the old and new versions, switching to the new version requires restarting the notebook kernel. With \texttt{auto\_rerun} enabled, \texttt{davos} takes care of this automatically and will restart the kernel, re-run the notebook up through the current cell, and also re-queue any other cells that the user had queued before the restart was triggered. This second time the code is run (as well as any future times), \texttt{smuggle} statements that have already been executed once before will run (virutlally) as quickly as import statements, since the specified package versions are already installed. Additionally,
